### 2015-07-21 text/ms.Rmd

In fisheries science, the commonly used operational model for determining status and trajectory of exploited fish populations are stock assessments that incorporate all appropriate data (e.g. catches, size and age distributions, surveys, and tagging information) to quantify values such as the biomass of a stock that can produce maximum sustainable yield ($B/B_\mathrm{MSY}$). However, the broad range of data required to conduct stock assessments are not available for the majority of fish populations, including those of conservation concern and of economic interest to fisheries [@fao2014]. Therefore, a number of models have been proposed to assess $B/B_\mathrm{MSY}$ based on the limited data available for the majority of fish stocks: catches and a basic understanding of population productivity [e.g. @vasconcellos2005; @martell2013]. Recently, @rosenberg2014 investigated the performance of four data-limited models through a large-scale simulation experiment. The four models frequently disagreed about population status and any one model had relatively poor predictive ability on average (e.g. Fig. \ref{motivate}).

@@

We first developed and tested ensemble methods with a fully factorial simulated dataset [@rosenberg2014] with precisely known status. These published simulations included the following scenarios: three fish life histories: small pelagic, demersal, and large pelagic; three levels of how depleted biomass was at the start of the dataset compared to carrying capacity: 100%, 70%, and 40% of carrying capacity; and four harvest dynamics: a constant harvest rate, a harvest rate coupled with biomass to mimic an open-access single-species fishery, a scenario where harvest rate increases continuously, and a 'roller-coaster' scenario where the harvest rate increases and then decreases. Process noise (i.e. variability in population dynamics) was introduced to the models through two levels of multiplicative recruitment variability: $N(0, 0.2)$ and $N(0, 0.6)$, and was either uncorrelated or had a first-order autoregressive correlation of $0.6$. The simulation also included a scenario with multiplicative observation error around catch at $N(0, 0.2)$ and one without observation error.

@@

Combining predictions from multiple models via superensemble methods is also broadly useful in other subfields of ecology and fisheries science. Predictions about extinction risk are widely used at the national (e.g. the US Endangered Species Act and the Canadian Species at Risk Act) and international [e.g. the IUCN Red List, @iucn2015] levels. These risk assessments generally involve fitting regression models to outcomes for individual species along with predictors of species risk [e.g. @anderson2011a; @pinsky2011], or population dynamics models to data for individual species [e.g. @dfo2010]. Both types of models are prone to error caused by model-misspecification and therefore results are sensitive to decisions about model structure [@brooks2015]. Although there are options to account for potential model-misspecification in determination of species risk [e.g. coefficient averaging, @burnham2002; generalized modeling, @yeakel2011; or semi-parametric methods, @thorson2014], ensemble methods are a relatively simple way to combine predictions in a transparent manner. Beyond estimates of status and trajectory, ensemble methods could be used to increase the robustness of spatial predictions when designing networks of protected areas [@rassweiler2014], to forecast potential spatial shifts in species distribution given climate impacts [@harsch2014], or to evaluate the effect of conservation or environmental management actions when experimental or observational data are limited *(Walsh et al. 2012)*. <!-- insert reference here: Walsh, J.C., Wilson, K.A., Benshemesh, J., Possingham, H.P., 2012. Unexpected outcomes of invasive predator control: the importance of evaluating conservation management actions. Animal Conservation 15, 319–328. -->

<!--

>>>>

In fisheries science, the commonly used operational model for determining status and trajectory of exploited fish populations are stock assessments that incorporate all appropriate data (e.g. catches, size and age distributions, surveys, and tagging information) to quantify values such as the biomass of a stock that can produce maximum sustainable yield ($B/B_\mathrm{MSY}$). However, the broad range of data required to conduct these stock assessments are not available for the majority of fish populations, including those of conservation concern and of economic interest to fisheries [@fao2014]. Therefore, a number of models have been proposed to assess $B/B_\mathrm{MSY}$ based on the limited data available for the majority of fish stocks: catches and a basic understanding of population productivity [e.g. @vasconcellos2005; @martell2013]. Recently, @rosenberg2014 investigated the performance of four data-limited models through a large-scale simulation experiment. The four models frequently disagreed about population status and any one model had relatively poor predictive ability on average (e.g. Fig. \ref{motivate}).

@@

We first developed and tested ensemble methods with a fully factorial simulated dataset [@rosenberg2014] with precisely known status. These published simulations included the following scenarios: three fish life histories: small pelagic, demersal, and large pelagic; three levels of how depleted biomass was at the start of the dataset compared to carrying capacity: 100%, 70%, and 40% of carrying capacity; and four harvest dynamics: a constant harvest rate, a harvest rate coupled with biomass to mimic an open-access single-species fishery, a scenario where harvest rate increases continuously, and a 'roller-coaster' scenario where the harvest rate increases and then decreases. Process noise (i.e. variability in population dynamics) was introduced to the models through two levels of multiplicative recruitment variability: $N(0, 0.2)$ and $N(0, 0.6)$, and was either uncorrelated or had a first-order autoregressive correlation of $0.6$. The simulation also included a scenario with multiplicative observation error around catch at $N(0, 0.2)$ and one without observation error. 
<!--TODO: make last sentence clearer-->

@@

Combining predictions from multiple models via superensemble methods is also broadly useful in other subfields of ecology and fisheries science. Predictions about extinction risk are widely used at the national (e.g. the US Endangered Species Act and the Canadian Species at Risk Act) and international [e.g. the IUCN Red List, @iucn2015] levels. These risk assessments generally involve fitting regression models to outcomes for individual species along with predictors of species risk [e.g. @anderson2011a; @pinsky2011], or population dynamics models to data for individual species [e.g. @dfo2010]. Both types of models are prone to error caused by model-misspecification and therefore results are sensitive to decisions about model structure [@brooks2015]. Although there are options to account for potential model-misspecification in determination of species risk [e.g. coefficient averaging, @burnham2002; generalized modeling, @yeakel2011; or semi-parametric methods, @thorson2014], ensemble methods are a relatively simple way to combine predictions in a transparent manner. Beyond estimates of status and trajectory, ensemble methods could be used to increase the robustness of spatial predictions when designing networks of protected areas [@rassweiler2014], to forecast potential spatial shifts in species distribution given climate impacts [@harsch2014], or to evaluate the effect of conservation management actions when experimental or observational data are limited [@walsh2012].
<!--
